---
title: "Week 02: COVID Visualization Activity"
date: "September, 11 2023"
author: "Aiganysh Ulanova"
---

Today, we'll be working with a data set related to COVID.
This data is based on data from the the [COVID Tracking Project](https://covidtracking.com/).
I cleaned up this data and also added total populations from the 2020 for each of the relevant categories.
Note, due to differences in the way race and ethnicity are encoded in the census as compared to the the COVID Tracking Project, the population counts for LatinX may be somewhat inaccurate.

```{r load_data}
library(tidyverse)
load("CRDT Data.RData")
ls()
```

I've include 4 different data sets.
They all have the same data but have it represented in different ways.
Try using the different data sets and see which ones are good for making which plots.


```{r first_plot}
ggplot(covid_data_count, aes(x = date, y = Cases)) + geom_point()
```



If you want to only look at a specific state, you can do it like this.
For now, see what you can do just using `ggplot`.

```{r}
covid_data_count |> 
  filter(state == "MA") |> 
  ggplot(aes(x = date, y = Cases, color = race)) + geom_line()
```

1. Write code for a plot and describe what you observe?
```{r}
# Attempt 1: Plotting deaths against date, colored by race, only for the state of Massachusetts

covid_data_count |> 
  filter(state == "MA") |> 
  ggplot(aes(x = date, y = Deaths, color = race)) + geom_line()

```
For this particular graph above, I see how the "Deaths" count is the highest in abundance and grows at the quickest rate for the race "White", which makes sense as proportionally, this race has the largest population in Massachusetts. The abundance and growth rate for the "Deaths" count is very similar for all other races (excluding "White)". It is interesting to note how for all races (although you can see this the most clearly for the "White" race), there is a huge increase in the rate for the "Deaths" count around the May - June area. You can clearly see the straight "up" slope for the "White" race (a rapid spike). This is most likely when the height of the pandemic was happening and when the "Cases" count grew rapidly as well (although this needs to be verified by looking more closely at the data). 

```{r}
# Attempt #2: Plot of deaths and cases simultaneously for "White" race in MA
covid_data_count |> 
  filter(state == "MA") |> 
  filter(race == "White") |>
  ggplot(aes(x = date)) + geom_line(aes(y=Cases), color= "black") + geom_line(aes(y=Deaths), color= "red") + labs(y="count", title = "Cases and Deaths count for the White population in MA") 

```
```
2. Are there any conclusions you can draw?
It's quite challenging to draw conclusions from a singular graph. But we can see how in the time frame from Oct - February, the rate of growth for cases grew at a much quicker rate than the rate of deaths. The rates were fairly comparable from April- May, but then they started to drastically differ.

3. What were you not able to do due to not having the R knowledge?
I wanted to plot barcharts where each bar would represent a race and the y-axis would be a proportion of the population that has gotten COVID or the proportion of the population that has died from COVID-19. However, at the time of writing I do not have the R knowledge for this. 

4. What other data would be useful to better understand this data?
Another data that would be useful is to look at the data for recoveries. Having the recovery count would be helpful to understand how the rate of recoveries compares to rate of growth for cases and deaths.

